---
title: "Data Exploration"
date: "2022-11-08 <br> Updated: `r Sys.Date()`"
format: pdf
editor: 
  markdown: 
    wrap: sentence
---

# ⭐️Overview

This file seeks to perform initial exploration of the NHANES data.

## 🕒 Change Log

**2023-11-08**, file initialized

# 📦 Load packages & Functions

```{r, message=FALSE, warning=FALSE}
library(here, warn.conflicts = FALSE)
library(tidyverse, warn.conflicts = FALSE)
library(mice, warn.conflicts = FALSE)
```

## 📥 Import the Data

We imported our cleaned NHANES data set.

```{r}
nhanes_path <- here('data', 'combined_data.rds')
nhanes_data <- readRDS(nhanes_path)

# Print a message for when this file is being sourced
cat(
  paste0(Sys.Date(), ":"),
  "NHANES data imported with", nrow(nhanes_data), "rows and", 
  ncol(nhanes_data), "columns. \n",
  "Data last modified: ", as.character(file.info(nhanes_path)$mtime), "\n"
)

# 2023-11-08: NHANES data imported with 7122 rows and 102 columns. 
#  Data last modified:  2023-11-08 11:39:26 
```

# Miss-Pattern

## All Subjects

There was at least one missing value in 86 of our 102 variables.

```{r}
contains_na <- nhanes_data |>
  select(where(~sum(is.na(.)) > 0)) |>
  select(-c(
    rx_statin, rx_pcks9, rx_fibrate, rx_sequestrant, rx_supplement,
    rx_other, rx_non_lipid, rx_glucose_lowering
    )) |>
  names()

length(contains_na)
#86
```

### Administrative Variables

None of the administratve variables had any missingness.

```{r}
length(
  nhanes_data |>
    select(
      id, drop_flag, collection_cycle, collection_lang
    ) |>
    select(where(~sum(is.na(.)) > 0))|>
    names()
)
# 0
```
### Inclusion/Exclusion Variables

7 of the 8 inclusion/exclusion variables demonstrated missingness.

```{r}
length(
  nhanes_data |>
    select(
      cholesterol_rx_ever, adherence, cancer_ever, rx_glucose_lowering,
      cholesterol_ever, cholesterol_check_ever, cholesterol_check_last,
      age
    ) |>
    select(where(~sum(is.na(.)) > 0))|>
    names()
)
# 7
```
We examined this pattern in greater detail. Age was the least missing. 

```{r}
md.pattern(nhanes_data |>
    select(
      cholesterol_rx_ever, adherence, cancer_ever, rx_glucose_lowering,
      cholesterol_ever, cholesterol_check_ever, cholesterol_check_last,
      age
    ), rotate.names = TRUE)
```
### Non-Insurance SES

7 of the 11 non-insurance SES variables demonstrated missingness.

```{r}
length(
  nhanes_data |>
    select(
      age, sex, race_5cat, race_6cat, marital_status, educ_level, 
      occupation_type_week, accult_birth_country, accult_time_usa, 
      income_ratio, income_cat
    ) |>
    select(where(~sum(is.na(.)) > 0))|>
    names()
)
# 7
```

We examined this pattern in greater detail. Marital Status, Income variables, and acculturation time were the most missing.

```{r}
md.pattern(nhanes_data |>
    select(
      age, sex, race_5cat, race_6cat, marital_status, educ_level, 
      occupation_type_week, accult_birth_country, accult_time_usa, 
      income_ratio, income_cat
    ), rotate.names = TRUE)
```
### Insurance Status Variables

All of the 14 insurance variables demonstrated missingness.

```{r}
length(
  nhanes_data |>
    select(
      ins_dichot_original, ins_classif, ins_type, ins_any_gap, ins_private, 
      ins_govt, ins_none, ins_rx, ins_govt_medicare, ins_govt_medigap, 
      ins_govt_medicaid, ins_govt_chip, ins_govt_state, ins_govt_other
    ) |>
    select(where(~sum(is.na(.)) > 0))|>
    names()
)
# 14
```

We examined this pattern in greater detail. The binary flags were the most frequently missing.

```{r}
md.pattern(nhanes_data |>
    select(
      ins_dichot_original, ins_classif, ins_type, ins_any_gap, ins_private, 
      ins_govt, ins_none, ins_rx, ins_govt_medicare, ins_govt_medigap, 
      ins_govt_medicaid, ins_govt_chip, ins_govt_state, ins_govt_other
    ), rotate.names = TRUE)
```
### Comorbid CHD Variables

All of the 11 comorbid CHD variables demonstrated missingness.

```{r}
length(
  nhanes_data |>
    select(
      cvd_hx, cvd_chf_ever, cvd_chf_age, cvd_cad_ever, cvd_cad_age, 
      cvd_angina_ever, cvd_angina_age, cvd_mi_ever, cvd_mi_age, cvd_cva_ever, 
      cvd_cva_age
    ) |>
    select(where(~sum(is.na(.)) > 0))|>
    names()
)
# 11
```

We examined this pattern in greater detail. The age variables were the most frequently missing.

```{r}
md.pattern(nhanes_data |>
    select(
      cvd_hx, cvd_chf_ever, cvd_chf_age, cvd_cad_ever, cvd_cad_age, 
      cvd_angina_ever, cvd_angina_age, cvd_mi_ever, cvd_mi_age, cvd_cva_ever, 
      cvd_cva_age
    ), rotate.names = TRUE)
```
### Substance Use Variables

All of the 5 substance use variables demonstrated missingness.

```{r}
length(
  nhanes_data |>
    select(
      alc_ever, alc_freq, alc_per_day, alc_heavy, smoking_hx
    ) |>
    select(where(~sum(is.na(.)) > 0))|>
    names()
)
# 5
```
We examined this pattern in greater detail. Alcohol per day was the most common source variable that was missing.

```{r}
md.pattern(nhanes_data |>
    select(
      alc_ever, alc_freq, alc_per_day, alc_heavy, smoking_hx
    ), rotate.names = TRUE)
```
### Physical Activity Variables

All of the 5 physical activity variables demonstrated missingness.

```{r}
length(
  nhanes_data |>
    select(
      active_mod_work_dur, active_transport_dur, active_mod_rec_dur, 
      active_total, active_vig_dur
    ) |>
    select(where(~sum(is.na(.)) > 0))|>
    names()
)
# 5
```

We examined this pattern in greater detail. Vigorous activity and transport activity durations were the most frequently missing.

```{r}
md.pattern(nhanes_data |>
    select(
      active_mod_work_dur, active_transport_dur, active_mod_rec_dur, 
      active_total, active_vig_dur
    ), rotate.names = TRUE)
```
### BMI and Weight Perception Variables

All of the 4 bmi and weight perception variables demonstrated missingness.

```{r}
length(
  nhanes_data |>
    select(
      bmi, bmi_cat, weight_self, weight_self_wish
    ) |>
    select(where(~sum(is.na(.)) > 0))|>
    names()
)
# 4
```
We examined this pattern in greater detail. BMI measurement was the most frequently missing value.

```{r}
md.pattern(nhanes_data |>
    select(
      bmi, bmi_cat, weight_self, weight_self_wish
    ), rotate.names = TRUE)
```
### Physician Weight Notification Variables

All of the 4 physician notification of weight needs variables demonstrated missingness.

```{r}
length(
  nhanes_data |>
    select(
      weight_need_ever, weight_need_exercise, weight_need_salt, 
      weight_need_fat
    ) |>
    select(where(~sum(is.na(.)) > 0))|>
    names()
)
# 4
```

We examined this pattern in greater detail. No particular variable was significantly more missing than any other.

```{r}
md.pattern(nhanes_data |>
    select(
      weight_need_ever, weight_need_exercise, weight_need_salt, 
      weight_need_fat
    ), rotate.names = TRUE)
```
### Current Weight Loss Activity Variables

All of the 4 current weight loss activity variables demonstrated missingness.

```{r}
length(
  nhanes_data |>
    select(
      weight_act_now, weight_act_now_exercise, weight_act_now_salt,
      weight_act_now_fat
    ) |>
    select(where(~sum(is.na(.)) > 0))|>
    names()
)
# 4
```

We examined this pattern in greater detail. No variable appeared to be significantly more missing than any others

```{r}
md.pattern(nhanes_data |>
    select(
      weight_act_now, weight_act_now_exercise, weight_act_now_salt,
      weight_act_now_fat
    ), rotate.names = TRUE)
```
### Past Year Weight Loss Activity Variables
All of the 22 past year weight-loss action variables demonstrated missingness.

```{r}
length(
  nhanes_data |>
    select(
      weight_act_year, weight_act_year_cut, weight_act_year_calories,
      weight_act_year_fat, weight_act_year_exercise, weight_act_year_skip,
      weight_act_year_product, weight_act_year_liquid, 
      weight_act_year_program, weight_act_year_rx, weight_act_year_nonrx,
      weight_act_year_purge, weight_act_year_water, weight_act_year_diet,
      weight_act_year_carbs, weight_act_year_smoke, weight_act_year_veg,
      weight_act_year_habits, weight_act_year_sweets, weight_act_year_junk,
      weight_act_year_surgery, weight_act_year_other
    ) |>
    select(where(~sum(is.na(.)) > 0))|>
    names()
)
# 22
```

We examined this pattern in greater detail. No one variable appeared to be significantly more missing than any other.

```{r}
md.pattern(nhanes_data |>
    select(
      weight_act_year, weight_act_year_cut, weight_act_year_calories,
      weight_act_year_fat, weight_act_year_exercise, weight_act_year_skip,
      weight_act_year_product, weight_act_year_liquid, 
      weight_act_year_program, weight_act_year_rx, weight_act_year_nonrx,
      weight_act_year_purge, weight_act_year_water, weight_act_year_diet,
      weight_act_year_carbs, weight_act_year_smoke, weight_act_year_veg,
      weight_act_year_habits, weight_act_year_sweets, weight_act_year_junk,
      weight_act_year_surgery, weight_act_year_other
    ), rotate.names = TRUE)
```
### RX Data Variables

All of the 11 prescription use variables demonstrated missingness.

```{r}
length(
  nhanes_data |>
    select(
      rx_use, rx_lipid, rx_count_lipid, rx_count_total, rx_statin, rx_pcks9, 
      rx_fibrate, rx_sequestrant, rx_supplement, rx_other, rx_non_lipid
    ) |>
    select(where(~sum(is.na(.)) > 0))|>
    names()
)
# 11
```

We examined this pattern in greater detail. The classification flags were the most frequently missing variables.

```{r}
md.pattern(nhanes_data |>
    select(
      rx_use, rx_lipid, rx_count_lipid, rx_count_total, rx_statin, rx_pcks9, 
      rx_fibrate, rx_sequestrant, rx_supplement, rx_other, rx_non_lipid
    ), rotate.names = TRUE)
```

## Included Subjects

We isolated only the subjects flagged for inclusion

```{r}
included_data <- nhanes_data |>
  filter(!drop_flag)
```

There was at least one missing value in 73 of our 102 variables.

```{r}
contains_na <- included_data |>
  select(where(~sum(is.na(.)) > 0)) |>
  select(-c(
    rx_statin, rx_pcks9, rx_fibrate, rx_sequestrant, rx_supplement,
    rx_other, rx_non_lipid, rx_glucose_lowering
    )) |>
  names()

length(contains_na)
#73
```

### Administrative Variables

None of the administratve variables had any missingness.

```{r}
length(
  included_data |>
    select(
      id, drop_flag, collection_cycle, collection_lang
    ) |>
    select(where(~sum(is.na(.)) > 0))|>
    names()
)
# 0
```

### Inclusion/Exclusion Variables

7 of the 8 inclusion/exclusion variables demonstrated missingness.

```{r}
length(
  included_data |>
    select(
      cholesterol_rx_ever, adherence, cancer_ever, rx_glucose_lowering,
      cholesterol_ever, cholesterol_check_ever, cholesterol_check_last,
      age
    ) |>
    select(where(~sum(is.na(.)) > 0))|>
    names()
)
# 5
```
We examined this pattern in greater detail. The binary flag for glucose lowering medication was the most frequently missing, followed by endorsement of cholesterol ever being checked.

```{r}
md.pattern(included_data |>
    select(
      cholesterol_rx_ever, adherence, cancer_ever, rx_glucose_lowering,
      cholesterol_ever, cholesterol_check_ever, cholesterol_check_last,
      age
    ), rotate.names = TRUE)
```

### Non-Insurance SES

2 of the 11 non-insurance SES variables demonstrated missingness.

```{r}
length(
  included_data |>
    select(
      age, sex, race_5cat, race_6cat, marital_status, educ_level, 
      occupation_type_week, accult_birth_country, accult_time_usa, 
      income_ratio, income_cat
    ) |>
    select(where(~sum(is.na(.)) > 0))|>
    names()
)
# 2
```

We examined this pattern in greater detail. Acculturation time was the most frequently missing.

```{r}
md.pattern(included_data |>
    select(
      age, sex, race_5cat, race_6cat, marital_status, educ_level, 
      occupation_type_week, accult_birth_country, accult_time_usa, 
      income_ratio, income_cat
    ), rotate.names = TRUE)
```

### Insurance Status Variables

13 of the 14 insurance variables demonstrated missingness.

```{r}
length(
  included_data |>
    select(
      ins_dichot_original, ins_classif, ins_type, ins_any_gap, ins_private, 
      ins_govt, ins_none, ins_rx, ins_govt_medicare, ins_govt_medigap, 
      ins_govt_medicaid, ins_govt_chip, ins_govt_state, ins_govt_other
    ) |>
    select(where(~sum(is.na(.)) > 0))|>
    names()
)
# 14
```

We examined this pattern in greater detail. The binary flags were the most frequently missing.

```{r}
md.pattern(included_data |>
    select(
      ins_dichot_original, ins_classif, ins_type, ins_any_gap, ins_private, 
      ins_govt, ins_none, ins_rx, ins_govt_medicare, ins_govt_medigap, 
      ins_govt_medicaid, ins_govt_chip, ins_govt_state, ins_govt_other
    ), rotate.names = TRUE)
```

### Comorbid CHD Variables

All of the 11 comorbid CHD variables demonstrated missingness.

```{r}
length(
  included_data |>
    select(
      cvd_hx, cvd_chf_ever, cvd_chf_age, cvd_cad_ever, cvd_cad_age, 
      cvd_angina_ever, cvd_angina_age, cvd_mi_ever, cvd_mi_age, cvd_cva_ever, 
      cvd_cva_age
    ) |>
    select(where(~sum(is.na(.)) > 0))|>
    names()
)
# 11
```

We examined this pattern in greater detail. The age variables were the most frequently missing.

```{r}
md.pattern(included_data |>
    select(
      cvd_hx, cvd_chf_ever, cvd_chf_age, cvd_cad_ever, cvd_cad_age, 
      cvd_angina_ever, cvd_angina_age, cvd_mi_ever, cvd_mi_age, cvd_cva_ever, 
      cvd_cva_age
    ), rotate.names = TRUE)
```

### Substance Use Variables

4 of the 5 substance use variables demonstrated missingness.

```{r}
length(
  included_data |>
    select(
      alc_ever, alc_freq, alc_per_day, alc_heavy, smoking_hx
    ) |>
    select(where(~sum(is.na(.)) > 0))|>
    names()
)
# 4
```
We examined this pattern in greater detail. Alcohol per day was the most common source variable that was missing.

```{r}
md.pattern(included_data |>
    select(
      alc_ever, alc_freq, alc_per_day, alc_heavy, smoking_hx
    ), rotate.names = TRUE)
```

### Physical Activity Variables

All of the 5 physical activity variables demonstrated missingness.

```{r}
length(
  included_data |>
    select(
      active_mod_work_dur, active_transport_dur, active_mod_rec_dur, 
      active_total, active_vig_dur
    ) |>
    select(where(~sum(is.na(.)) > 0))|>
    names()
)
# 5
```

We examined this pattern in greater detail. Vigorous activity and transport activity durations were the most frequently missing.

```{r}
md.pattern(included_data |>
    select(
      active_mod_work_dur, active_transport_dur, active_mod_rec_dur, 
      active_total, active_vig_dur
    ), rotate.names = TRUE)
```

### BMI and Weight Perception Variables

3 of the 4 bmi and weight perception variables demonstrated missingness.

```{r}
length(
  included_data |>
    select(
      bmi, bmi_cat, weight_self, weight_self_wish
    ) |>
    select(where(~sum(is.na(.)) > 0))|>
    names()
)
# 3
```
We examined this pattern in greater detail. BMI measurement was the most frequently missing value.

```{r}
md.pattern(included_data |>
    select(
      bmi, bmi_cat, weight_self, weight_self_wish
    ), rotate.names = TRUE)
```

### Physician Weight Notification Variables

2 of the 4 physician notification of weight needs variables demonstrated missingness.

```{r}
length(
  included_data |>
    select(
      weight_need_ever, weight_need_exercise, weight_need_salt, 
      weight_need_fat
    ) |>
    select(where(~sum(is.na(.)) > 0))|>
    names()
)
# 2
```

We examined this pattern in greater detail. No particular variable was significantly more missing than any other.

```{r}
md.pattern(included_data |>
    select(
      weight_need_ever, weight_need_exercise, weight_need_salt, 
      weight_need_fat
    ), rotate.names = TRUE)
```

### Current Weight Loss Activity Variables

3 of the 4 current weight loss activity variables demonstrated missingness.

```{r}
length(
  included_data |>
    select(
      weight_act_now, weight_act_now_exercise, weight_act_now_salt,
      weight_act_now_fat
    ) |>
    select(where(~sum(is.na(.)) > 0))|>
    names()
)
# 3
```

We examined this pattern in greater detail. No variable appeared to be significantly more missing than any others

```{r}
md.pattern(included_data |>
    select(
      weight_act_now, weight_act_now_exercise, weight_act_now_salt,
      weight_act_now_fat
    ), rotate.names = TRUE)
```

### Past Year Weight Loss Activity Variables

22 of the 22 past year weight-loss action variables demonstrated missingness.

```{r}
length(
  included_data |>
    select(
      weight_act_year, weight_act_year_cut, weight_act_year_calories,
      weight_act_year_fat, weight_act_year_exercise, weight_act_year_skip,
      weight_act_year_product, weight_act_year_liquid, 
      weight_act_year_program, weight_act_year_rx, weight_act_year_nonrx,
      weight_act_year_purge, weight_act_year_water, weight_act_year_diet,
      weight_act_year_carbs, weight_act_year_smoke, weight_act_year_veg,
      weight_act_year_habits, weight_act_year_sweets, weight_act_year_junk,
      weight_act_year_surgery, weight_act_year_other
    ) |>
    select(where(~sum(is.na(.)) > 0))|>
    names()
)
# 22
```

We examined this pattern in greater detail. No one variable appeared to be significantly more missing than any other.

```{r}
md.pattern(included_data |>
    select(
      weight_act_year, weight_act_year_cut, weight_act_year_calories,
      weight_act_year_fat, weight_act_year_exercise, weight_act_year_skip,
      weight_act_year_product, weight_act_year_liquid, 
      weight_act_year_program, weight_act_year_rx, weight_act_year_nonrx,
      weight_act_year_purge, weight_act_year_water, weight_act_year_diet,
      weight_act_year_carbs, weight_act_year_smoke, weight_act_year_veg,
      weight_act_year_habits, weight_act_year_sweets, weight_act_year_junk,
      weight_act_year_surgery, weight_act_year_other
    ), rotate.names = TRUE)
```

### RX Data Variables

All of the 11 prescription use variables demonstrated missingness.

```{r}
length(
  included_data |>
    select(
      rx_use, rx_lipid, rx_count_lipid, rx_count_total, rx_statin, rx_pcks9, 
      rx_fibrate, rx_sequestrant, rx_supplement, rx_other, rx_non_lipid
    ) |>
    select(where(~sum(is.na(.)) > 0))|>
    names()
)
# 11
```

We examined this pattern in greater detail. The classification flags were the most frequently missing variables.

```{r}
md.pattern(included_data |>
    select(
      rx_use, rx_lipid, rx_count_lipid, rx_count_total, rx_statin, rx_pcks9, 
      rx_fibrate, rx_sequestrant, rx_supplement, rx_other, rx_non_lipid
    ), rotate.names = TRUE)
```

