---
title: "Summary Statistics Tables"
date: "2022-11-25 <br> Updated: `r Sys.Date()`"
format: pdf
editor: 
  markdown: 
    wrap: sentence
---

# ⭐️Overview

This file seeks to generate summary tables.

# 📦 Load packages & Functions

```{r, message=FALSE, warning=FALSE}
library(here, warn.conflicts = FALSE)
library(tidyverse, warn.conflicts = FALSE)
library(mice, warn.conflicts = FALSE)
```

```{r}
source(here("R", "comparative_stats_table.R"))
```

## 📥 Import the Data

We imported our cleaned NHANES data set.

```{r}
nhanes_path <- here('data', 'combined_data.rds')
nhanes_data <- readRDS(nhanes_path)

# Print a message for when this file is being sourced
cat(
  paste0(Sys.Date(), ":"),
  "NHANES data imported with", nrow(nhanes_data), "rows and", 
  ncol(nhanes_data), "columns. \n",
  "Data last modified: ", as.character(file.info(nhanes_path)$mtime), "\n"
)

# 2023-11-25: NHANES data imported with 7122 rows and 104 columns. 
#  Data last modified:  2023-11-25 15:10:57 
```

We also imported our cleaned subset of included subjects

```{r}
included_path <- here('data', 'subset_1168.rds')
included_data <- readRDS(included_path)

# Print a message for when this file is being sourced
cat(
  paste0(Sys.Date(), ":"),
  "Included subject subset data imported with", nrow(included_data), "rows and", 
  ncol(included_data), "columns. \n",
  "Data last modified: ", as.character(file.info(included_path)$mtime), "\n"
)

# 2023-11-25: Included subject subset data imported with 1168 rows and 104 columns. 
#  Data last modified:  2023-11-25 15:10:57 
```

# Reverse Ordering of Factors

# Variable Clusters

We identified clusters of variables for our tables.

```{r}
demographics <- c(
  'age', 'sex', 'race_ethn_5cat', 'marital_status', 'educ_level', 
  'income_ratio', 'income_cat'
  )

insurance_variables <- c('ins_dichot', 'ins_type', 'ins_rx')

health_vars <- c(
  'bmi', 'bmi_cat', 'alc_heavy', 'smoking_hx', 'cvd_hx', 'cvd_chf_ever', 
  'cvd_cad_ever', 'cvd_angina_ever', 'cvd_mi_ever', 'cvd_cva_ever'
  )

rx_vars <- c(
  'rx_use', 'rx_lipid', 'rx_count_lipid', 'rx_count_total', 'rx_statin', 
  'rx_pcks9', 'rx_fibrate', 'rx_sequestrant', 'rx_supplement', 'rx_other'
  )

desired_vars <- c(demographics, insurance_variables, health_vars, rx_vars)

cat_vars <- c(
  'sex', 'race_ethn_5cat', 'marital_status', 'educ_level', 'income_cat',
  'bmi_cat', 'alc_heavy', 'smoking_hx', 'cvd_hx', 'cvd_chf_ever', 
  'cvd_cad_ever', 'cvd_angina_ever', 'cvd_mi_ever', 'cvd_cva_ever',
  'ins_dichot', 'ins_type', 'ins_rx',
  'rx_use', 'rx_lipid', 'rx_statin', 'rx_pcks9', 'rx_fibrate', 
  'rx_sequestrant', 'rx_supplement', 'rx_other'
  )
```

# All Subjects

We extracted the desired columns from the complete subject data set

```{r}
complete_data <- complete_data |>
  select(all_of(desired_vars), adherence)
```

We generated our comparative statistics table. 

```{r}
comparative_stats_table(
  complete_data,
  dichot_var = 'adherence',
  cont_mean_cols = c('age'),
  cont_median_cols = c('income_ratio', 'bmi', 'rx_count_total', 'rx_count_lipid'),
  cat_cols = cat_vars,
  var_order = desired_vars,
  docx_path = here('tables', 'Comparative Table - All Subjects by Adherence.docx')
)
```

# Included Subjects

We extracted the desired columns from the included subject data set

```{r}
included_data <- included_data |>
  select(all_of(desired_vars), adherence, id)
```

We generated our comparative statistics table using Median for Age.

```{r}
comparative_stats_table(
  included_data,
  dichot_var = 'adherence',
  # ID used for mean since no variables are normally distributed in this case.
  cont_mean_cols = c('id'),
  cont_median_cols = c('age', 'income_ratio', 'bmi', 'rx_count_total', 
                       'rx_count_lipid'),
  cat_cols = cat_vars,
  var_order = desired_vars,
  docx_path = here('tables', 'Comparative Table - Included Subjects by Adherence - Age Median.docx')
)
```

We generated our comparative statistics table using Mean for Age.

```{r}
comparative_stats_table(
  included_data,
  dichot_var = 'adherence',
  # ID used for mean since no variables are normally distributed in this case.
  cont_mean_cols = c('age'),
  cont_median_cols = c('income_ratio', 'bmi', 'rx_count_total', 
                       'rx_count_lipid'),
  cat_cols = cat_vars,
  var_order = desired_vars,
  docx_path = here('tables', 'Comparative Table - Included Subjects by Adherence - Age Mean.docx')
)
```